### 2008-12-18 talk/ecoop2009/tlc.tex
fix some things in the tlc section

In this section, we will briefly describe \emph{TLC}, a simple dynamic
language that we developed to exercise our JIT compiler generator.  As most of
dynamic languages around, \emph{TLC} is implemented through a virtual machine
that interprets a custom bytecode. Since our main interest is in the runtime
performance of the VM, we did not implement the parser nor the bytecode
compiler, but only the VM itself.

@@

The virtual machine is stack-based, and provides several operations:

\begin{itemize}
\item \textbf{Stack manipulation}: standard operations to manipulate the
  stack, such as \lstinline{PUSH}, \lstinline{POP}, \lstinline{SWAP}, etc.
\item \textbf{Flow control} to do conditional and unconditional jumps.
\item \textbf{Arithmetic}: numerical operations on integers, like
  \lstinline{ADD}, \lstinline{SUB}, etc.
\item \textbf{Comparisons} like \lstinline{EQ}, \lstinline{LT},
  \lstinline{GT}, etc.
\item \textbf{Object-oriented}: operations on objects: \lstinline{NEW},
  \lstinline{GETATTR}, \lstinline{SETATTR}, \lstinline{SEND}.
\item \textbf{List operations}: \lstinline{CONS}, \lstinline{CAR},
  \lstinline{CDR}.
\end{itemize}

@@

\subsection{TLC features}
\label{sec:tlc-features}

Despite being very simple and minimalistic, \lstinline{TLC} is a good
candidate as a language to test our JIT generator, as it has some of the
features that makes most of current dynamic languages so slow:

@@

\item \textbf{Stack based VM}: this kind of VM requires all the operands to be
  on top of the evaluation stack.  As a consequence programs spend a lot of
  time pushing and popping values to/from the stack, or doing other stack
  related operations.  However, thanks to its simplicity this is still the
  most common and preferred way to implement VMs.

@@

\begin{lstlisting}
main:             # stack: []
    PUSHARG       #        [n]
    PUSH 0        #        [n, 0]
    LT            #        [n<0]
    BR_COND neg

>>>>

In this section, we will briefly describe \emph{TLC}, a simple dynamic
language that we developed to exercise our JIT compiler generator and that will
be used as a running example in this paper. The design goal of the language is
to be very simple (the interpreter of the full language consists of about 600
lines of RPython code) but to still have the typical properties of dynamic
languages that make them hard to compile. \emph{TLC} is implemented with a small
interpreter that interprets a custom bytecode instruction set. Since our main
interest is in the runtime performance of the interpreter, we did not implement
the parser nor the bytecode compiler, but only the interpreter itself.

@@

The interpreter for the language is stack-based and uses bytecode to represent
the program. It provides the following bytecode instructions:

\begin{itemize}
\item \textbf{Stack manipulation}: standard operations to manipulate the
  stack, such as \lstinline{POP}, \lstinline{PUSHARG}, \lstinline{SWAP}, etc.
\item \textbf{Flow control} to do conditional and unconditional jumps.
\cfbolz{what is the bytecode for conditional jumps?}
\item \textbf{Arithmetic}: numerical operations on integers, like
  \lstinline{ADD}, \lstinline{SUB}, etc.
\item \textbf{Comparisons} like \lstinline{EQ}, \lstinline{LT},
  \lstinline{GT}, etc.
\item \textbf{Object-oriented}: operations on objects: \lstinline{NEW},
  \lstinline{GETATTR}, \lstinline{SETATTR}, \lstinline{SEND}.
\item \textbf{List operations}: \lstinline{CONS}, \lstinline{CAR},
  \lstinline{CDR}.
\end{itemize}

@@

\subsection{TLC features}
\label{sec:tlc-features}
\cfbolz{calling this sections "features" is a bit obscure, since it is more
properties of the implementation}

Despite being very simple and minimalistic, \lstinline{TLC} is a good
candidate as a language to test our JIT generator, as it has some of the
properties that makes most of current dynamic languages so slow:

@@

\item \textbf{Stack based interpreter}: this kind of interpreter requires all the operands to be
  on top of the evaluation stack.  As a consequence programs spend a lot of
  time pushing and popping values to/from the stack, or doing other stack
  related operations.  However, thanks to its simplicity this is still the
  most common and preferred way to implement interpreters.

@@

\begin{lstlisting}
main:             # stack: []
    PUSHARG       #        [n]
    PUSH 0        #        [n, 0]
    LT            #        [0 or 1]
    BR_COND neg
